Progress Report

### 2020-04-29 documentation/systemModel/systemModel2.tex
Finished content for initial presentation

\subsection{Canceling Phases}
While deriving the fixed point equation for the AED, one step includes using additive free convolution to find the R-Transform for symmetrized singular value distributions using the property 
\begin{equation}
\tilde{R}_{\mathbf{H}_{Total}}(w) = \tilde{R}_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w) + \tilde{R}_{\mathbf{G}}(w).
\end{equation} In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w)$ and $R_{\mathbf{G}}(w)$, we use the property relating the symmetrized singular value distribution to the eigenvalue distribution.
\begin{equation}\label{svd_aed_property}
\tilde{G}_{B}(s) = sG_{BB^{\dagger}}(s^2).
\end{equation}
We begin by finding the stieltjes transform of the eigenvalue distributions of the individual component,
$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$, in order to use property \eqref{svd_aed_property} to find $\tilde{G}_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(s)$ (the symmetrized singular valued distribution).
\par
To find $G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ we define
\begin{equation}
C_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H
\end{equation}
and a rotated version 
\begin{equation}
\tilde{C}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
\end{equation}. 
These allow use of the rotational property of the S-Transform from \cite{muller2002asymptotic}
\begin{equation}\label{rotation_property}
S_{C_2}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_2}(\frac{z}{\chi_2}).
\end{equation}
in which $\chi_N$ is the aspect ratio of the channel matrices. In the given model $\chi_2 = \frac{S}{R}$ and
$\chi_2 = \frac{S}{R}$.
Applying the rotational property again to the term 
\begin{equation}
C_{1} = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
\end{equation}
we get 
\begin{equation}
\tilde{C}_{1} \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
\end{equation}
with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$.

From this we can conclude that for $R,S,T \rightarrow \infty $ the only impact of the IRS comes from the aspect ratio
$\chi_N$ but the actual phases of the IRS do not impact the final AED.
The results shown in \cite{zhang2019capacity} do not reflect the result only because 
the values $R = 4, S = 100 , T = 4$ are chosen so these asymptotic results do not apply. 
These results can be seem in

\section{Optimization}
For the system
\begin{equation}
		\mathbf{r} = \mathbf{H}_{Total}\mathbf{x},
	\end{equation}
	with  $R,S,T \rightarrow \infty $,
we now consider
how to choose the transmit covariance matrix $\mathbf{Q}_x$ and $\boldsymbol{\Phi}$ in order to optimize an objective.
First, We will optimize the precoder and IRS phases to maximize the channel capacity.
		    \begin{equation}
    \begin{array}{ll}
    \underset{\boldsymbol{\Phi}, \mathbf{Q}_x}{\text{maximize }}   & \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]\mathbf{Q}_x[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]^H|\right)\right]
    \\
    \mbox{subject to } & \|\phi_i\|^2 = 1
    \\
    & tr(\mathbf{Q}_x) \leq P_{\text{Total}}
%        \\
%    & \tilde{\mathbf{f}}_i \mathbf{h}_j = 0, \; i\neq j, \; \forall i,j \in N
    \end{array}
    \label{general_irs_opt}
    \end{equation}

@@

\subsection{Evaluating Impact of Cross-terms on AED}\label{cross_terms}
The polynomial describing the channel
covariance matrix given by 
\begin{equation}
\mathbf{H}_{Total}\mathbf{H}_{Total}^H = 
\underbrace{
\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}_{\text{cross terms}}
+ 
\underbrace{
\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}\mathbf{G}^H}_{\text{non-cross terms}}.
\end{equation} 
Here we wanted to see if the final AED is influenced by the cross terms. The case in which elements of the channel matrices are zero-mean with variance  $\frac{1}{N}$ is shown in \ref{cross}. In this case, the cross terms appear to have symmetry such that they will not influence the final AED.

@@

\subsection{Impact of Phases on total Capacity}\label{phase_impact}
It is hypothesized that in the asymptotic limit, if $\chi_n = 1$ the phases will not impact the capacity. In order to check this numerically I evaluated the capacity while holding the channel values constant for different values of the IRS phases. Here statistics are shown for 100 random realizations of the phases at the IRS. 
Phases, $e^{j\phi}$, are generated randomly over a uniform distribution $\phi \in [0,2\pi]$. This is clearly not rigorous but perhaps provides some insight. 
$T = R = S = 500$
\begin{itemize}
\item
Capacity variance:
 1.1512987368371963e-05
 \item
Capacity average:
 1.990962024050068
 \item
Capacity min:
 1.9818993465470727
 \item
Capacity max:
 1.9986482518130326
\end{itemize}
Similarly, the results for $T = 500 R = 500$ and $S = 1000$ are shown 
\begin{itemize}
\item
Capacity variance:
 1.730056273847992e-05
 \item 
Capacity average:
 1.9937729377849243
 \item
Capacity min:
 1.9855835490537985
 \item
Capacity max:
 2.0025075539557737
\end{itemize}

>>>>

\subsection{Canceling Phases}
While deriving the fixed point equation for the AED, one step includes using additive free convolution to find the R-Transform for symmetrized singular value distributions using the property 
\begin{equation}
\tilde{R}_{\mathbf{H}_{Total}}(w) = \tilde{R}_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w) + \tilde{R}_{\mathbf{G}}(w).
\end{equation} In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w)$ and $R_{\mathbf{G}}(w)$, we use the property relating the symmetrized singular value distribution to the eigenvalue distribution.
\begin{equation}\label{svd_aed_property}
\tilde{G}_{B}(s) = sG_{BB^{\dagger}}(s^2).
\end{equation}
We begin by finding the stieltjes transform of the eigenvalue distributions of the individual component,
$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$, in order to use property \eqref{svd_aed_property} to find $\tilde{G}_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(s)$ (the symmetrized singular valued distribution).
\par
To find $G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ we define
\begin{equation}
C_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H
\end{equation}
and a rotated version 
\begin{equation}
\tilde{C}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
\end{equation}. 
These allow use of the rotational property of the S-Transform from \cite{muller2002asymptotic}
\begin{equation}\label{rotation_property}
S_{C_2}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_2}(\frac{z}{\chi_2}).
\end{equation}
in which $\chi_N$ is the aspect ratio of the channel matrices. In the given model $\chi_2 = \frac{S}{R}$ and
$\chi_2 = \frac{S}{R}$.
Applying the rotational property again to the term 
\begin{equation}
C_{1} = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
\end{equation}
we get 
\begin{equation}
\tilde{C}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
\end{equation}
with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$.

From this we can conclude that for $R,S,T \rightarrow \infty $ the only impact of the IRS comes from the aspect ratio
$\chi_N$ but the actual phases of the IRS do not impact the final AED.
The results shown in \cite{zhang2019capacity} do not reflect that the phases play no role in capacity because 
the matrix dimensions chosen, $R = 4, S = 100 , T = 4$ are not applicable to asymptotic analysis. 
These results can be seen in 
\section{Optimization}
For the system
\begin{equation}
		\mathbf{r} = \mathbf{H}_{Total}\mathbf{x},
	\end{equation}
	with  $R,S,T \rightarrow \infty $,
we now consider
how to choose the transmit covariance matrix $\mathbf{Q}_x$ and $\boldsymbol{\Phi}$ in order to optimize an objective.
First, We will optimize the precoder and IRS phases to maximize the channel capacity.
		    \begin{equation}
    \begin{array}{ll}
    \underset{\boldsymbol{\Phi}, \mathbf{Q}_x}{\text{maximize }}   & \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]\mathbf{Q}_x[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]^H|\right)\right]
    \\
    \mbox{subject to } & \|\phi_i\|^2 = 1
    \\
    & tr(\mathbf{Q}_x) \leq P_{\text{Total}}
%        \\
%    & \tilde{\mathbf{f}}_i \mathbf{h}_j = 0, \; i\neq j, \; \forall i,j \in N
    \end{array}
    \label{general_irs_opt}
    \end{equation}

@@

\subsection{Evaluating Impact of Cross-terms on AED}\label{cross_terms}
The polynomial describing the channel
covariance matrix given by 
\begin{equation}
\mathbf{H}_{Total}\mathbf{H}_{Total}^H = 
\underbrace{
\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}\mathbf{G}^H}_{\text{cross terms}}
+ 
\underbrace{
\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}_{\text{non-cross terms}}.
\end{equation} 
Here we wanted to see if the final AED is influenced by the cross terms. The case in which elements of the channel matrices are zero-mean with variance  $\frac{1}{N}$ is shown in \ref{cross}. In this case, the cross terms appear to have symmetry such that they will not influence the final AED.

@@

\subsection{Impact of Phases on total Capacity}\label{phase_impact}
It is hypothesized that in the asymptotic limit, if $\chi_n = 1$ the phases will not impact the capacity. In order to check this numerically I evaluated the capacity while holding the channel values constant for different values of the IRS phases. Here statistics are shown for 100 random realizations of the phases at the IRS. 
Phases, $e^{j\phi}$, are generated randomly over a uniform distribution $\phi \in [0,2\pi]$. This is clearly not rigorous but perhaps provides some insight. 
$T = R = S = 500$
\begin{itemize}
\item
Capacity variance:
 1.1512987368371963e-05
 \item
Capacity average:
 1.990962024050068
 \item
Capacity min:
 1.9818993465470727
 \item
Capacity max:
 1.9986482518130326
\end{itemize}
Similarly, the results for $T = 500$, $ R = 500$ and $S = 1000$ are shown 
\begin{itemize}
\item
Capacity variance:
 1.730056273847992e-05
 \item 
Capacity average:
 1.9937729377849243
 \item
Capacity min:
 1.9855835490537985
 \item
Capacity max:
 2.0025075539557737
\end{itemize}
